### 2020-10-04 Neutrino.tex
tweak CIP

\paragraph{Example: three right-handed neutrinos}
Let us assume all the neutrinos are massive thanks to three right-handed neutrinos. Then $M\w L\wdiag$ is invertible and
\begin{equation}
 R:=-\ii (M\w H\wdiag)^{-1/2}\MD  U^{\text{Majorana}}\w{PMNS}  (M\w L^\text{diag})^{-1/2}
\quad\Longrightarrow\quad R^\TT R=1.
\end{equation}
Conversely, with a matrix $R$ satisfying $R^\TT R=1$, the ``Yukawa matrix'' is given by
\begin{equation}
 \MD = \ii \sqrt{M\w H^\text{diag}}R\sqrt{M\w L^\text{diag}}(U^{\text{Majorana}}\w{PMNS} )^\dagger.
\end{equation}
Now we successfully parameterized $Y_n$ by a ``complex orthogonal'' matrix $R$.

Such $R$ is given by\footnote{For $w\in\mathbb C$, $\sin z_1=w$ and $\cos z_2=w$ always have solutions $z_{1,2}\in\mathbb C$. Meanwhile, $\tan z=w$ has no solution if and only if $w=\pm \ii$. Then, using this fact, one first expresses $R_{i3}$ components by $\zeta\w{A,B,C}=\pm1$ and $\theta\w{A,B}\in\mathbb C$, restricting $0\le\Re\theta\w{A,B}\le\pi/2$ ($\Leftrightarrow\Re{\sin{\theta}}\ge0\land\Re\cos\theta\ge0$), and then gets an expression of $R$ with three angles and six signs. Five signs are absorbed by enlarging $\Re\theta$ and one sign remains, which is $\zeta$.}
\begin{equation}
 R=\pmat{
 \co{12} \co{13} & \si{12} \co{13} & \si{13} \\
 -\zeta \si{12} \co{23}-\co{12} \si{23} \si{13} & \zeta \co{12} \co{23} -\si{12} \si{23} \si{13} & \si{23} \co{13} \\
  \zeta \si{12} \si{23}-\co{12} \co{23} \si{13} & -\zeta\co{12} \si{23} -\si{12} \co{23} \si{13} & \co{23} \co{13}
},
\end{equation}
where $\co{12}\equiv \cos{\theta_{12}}$ etc.\ and
\begin{equation}
 \zeta=\pm1; \qquad
(\theta_{12},\theta_{23},\theta_{13})\in\mathbb C,\quad
|\Re\theta_{12}|\le\pi, \quad |\Re\theta_{23}|\le\pi,  \quad |\Re\theta_{13}|\le\frac{\pi}{2}.
\end{equation}
This $R$ satisfies $RR^\TT=1$, which however is not general (as in the next example).

>>>>

\paragraph{Example: three right-handed neutrinos}
Let us assume all the neutrinos are massive thanks to three right-handed neutrinos. Then $M\w L\wdiag$ is invertible and
\begin{equation}
 R:=-\ii (M\w H\wdiag)^{-1/2}\MD  U^{\text{Majorana}}\w{PMNS}  (M\w L^\text{diag})^{-1/2}
\quad\Longrightarrow\quad R^\TT R=1.
\end{equation}
Conversely, with a matrix $R$ satisfying $R^\TT R=1$, the ``Yukawa matrix'' is given by
\begin{equation}
 \MD = \ii \sqrt{M\w H^\text{diag}}R\sqrt{M\w L^\text{diag}}(U^{\text{Majorana}}\w{PMNS} )^\dagger.
\end{equation}
Now we successfully parameterized $Y_n$ by a ``complex orthogonal'' matrix $R$.
The extended PMNS matrix is given by
\begin{equation}
 \approx \pmat{
  U_1^* & \MD^\dagger \MN^{-1}U_2^*\\
  -\MN^{-1}\MD U_1^* & U_2^*
}
\approx
\pmat{
 U^{\text{Majorana}}\w{PMNS} & -\ii U^{\text{Majorana}}\w{PMNS} 
\sqrt{M\w L^\text{diag}}R^\dagger
(M\w H^\text{diag})^{-1/2}\\
  -\ii(M\w H^\text{diag})^{-1/2}R\sqrt{M\w L^\text{diag}}
&1
}.
\end{equation}

The parameter matrix $R$ is given by\footnote{For $w\in\mathbb C$, $\sin z_1=w$ and $\cos z_2=w$ always have solutions $z_{1,2}\in\mathbb C$. Meanwhile, $\tan z=w$ has no solution if and only if $w=\pm \ii$. Then, using this fact, one first expresses $R_{i3}$ components by $\zeta\w{A,B,C}=\pm1$ and $\theta\w{A,B}\in\mathbb C$, restricting $0\le\Re\theta\w{A,B}\le\pi/2$ ($\Leftrightarrow\Re{\sin{\theta}}\ge0\land\Re\cos\theta\ge0$), and then gets an expression of $R$ with three angles and six signs. Five signs are absorbed by enlarging $\Re\theta$ and one sign remains, which is $\zeta$.}
\begin{equation}
 R=\pmat{
 \co{12} \co{13} & \si{12} \co{13} & \si{13} \\
 -\zeta \si{12} \co{23}-\co{12} \si{23} \si{13} & \zeta \co{12} \co{23} -\si{12} \si{23} \si{13} & \si{23} \co{13} \\
  \zeta \si{12} \si{23}-\co{12} \co{23} \si{13} & -\zeta\co{12} \si{23} -\si{12} \co{23} \si{13} & \co{23} \co{13}
},
\end{equation}
where $\co{12}\equiv \cos{\theta_{12}}$ etc.\ and
\begin{equation}
 \zeta=\pm1; \qquad
(\theta_{12},\theta_{23},\theta_{13})\in\mathbb C,\quad
|\Re\theta_{12}|\le\pi, \quad |\Re\theta_{23}|\le\pi,  \quad |\Re\theta_{13}|\le\frac{\pi}{2}.
\end{equation}
This $R$ satisfies $RR^\TT=1$, which however is not general (as in the next example).
